---
title: "Covid19 vaccine study Final2024"
author: "Kaylie Fagoaga"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction
A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

##  Males
$H_0$:The covid921 vaccine is not effective in males 

$H_A$:The covid921 vaccine is effective in males 

###  Methods
treatment is a categorical variable   
infected is a categorical variable   
So we use is cat~cat helper file   
 
### Descriptive Results 


#### Graphical Results
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males, type="percent")

```

This barchart shows that males who took the vaccine were less likely to get infected with covid19 than the males who took the placebo.   


#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

The numerical summary shows that 74% of males who took the placebo got covid19 whereas 26% of males who took the vaccine got covid19.     

#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test has a p-valye of 0.00000000000000022.      
This means we reject the null hypothesis because the p-value of the chi-squared is less than 0.05.    
The Fisher exact test also had a p-value of 0.00000000000000022.      
The odds ratio shows that males who took the placebo are 2.8 more times likely to get covid19 than those who took the vaccine.      
This means we reject the null hypothesis because the p-value of the fisher exact test is less than 0.05.     

### Conclusion 
We conclude that the barchart shows males who took the vaccine were less likely to get infected with covid19 than the males who took the placebo.      
The Numerical summary shows that that 74% of males who took the placebo got covid19 whereas 26% of males who took the vaccine got covid19.   
The inferential results show that both the chi-squared test and Fisher exact test had a p-value of 0.00000000000000022.     
The odds ratio in the Fisher exact test shows that males who took the placebo are 2.8 more times likely to get covid19 than those who took the vaccine.  
This means we reject the null hypothesis that covid921 vaccine is not effective in males because both p-value's were less than 0.05.   

## Females
$H_0$:The covid921 vaccine is not effective in females 

$H_A$:The covid921 vaccine is effective in females

###  Methods
treatment is a categorical variable   
infected is a categorical variable   
So we use is cat~cat helper file  

### Descriptive Results 


#### Graphical Results
```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

The barchart shows that females who took the vaccine were less likely to get infected with covid19 than the females who took the placebo.   

#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected, data=females)
rowPerc(table1)
colPerc(table1)
```

The numerical summary shows that 65% of females who took the placebo got covid19 whereas 35% of females who took the vaccine got covid19.  


#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test has a p-valye of 0.00000001914.      
This means we reject the null hypothesis because the p-value of the chi-squared is less than 0.05.    
The Fisher exact test had a p-value of 0.00000001246.      
The odds ratio shows that feamles who took the placebo are 1.8 more times likely to get covid19 than those who took the vaccine.      
This means we reject the null hypothesis because the p-value of the fisher exact test is less than 0.05.    

### Conclusion 
We conclude that the barchart shows females who took the vaccine were less likely to get infected with covid19 than the females who took the placebo.         
The Numerical summary shows that that 65% of females who took the placebo got covid19 whereas 35% of females who took the vaccine got covid19.        
The inferential results show that the chi-squared test had a p-value of 0.00000001914.    
This means we reject the null hypothesis that covid921 vaccine is not effective in females because the p-value is less than 0.05.    
The Fisher exact test had a p-value of 0.00000001246.          
The odds ratio in the Fisher exact test shows that females who took the placebo are 1.8 more times likely to get covid19 than those who took the vaccine.       
This means we reject the null hypothesis that covid921 vaccine is not effective in females because the p-value is less than 0.05.    


## LGBTQ
$H_0$:The covid921 vaccine is not effective in the LGBTQ community 

$H_A$:The covid921 vaccine is effective in the LGBTQ community

###  Methods
treatment is a categorical variable   
infected is a categorical variable   
So we use is cat~cat helper file    
 
### Descriptive Results 


#### Graphical Results
```{r}
LGBTQ = filter(FinalData, LGBTQ=="gay")
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

The barchart shows that the LGBTQ who took the vaccine were less likely to get infected with covid19 than the LGBTQ who took the placebo.    

#### Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical summary shows that 97% of the LGBTQ who took the placebo got covid19 whereas 3% of the LGBTQ who took the vaccine got covid19.     


#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test has a p-valye of 0.000002266.              
This means we reject the null hypothesis because the p-value of the chi-squared is less than 0.05.       
The Fisher exact test had a p-value of 0.000000269.           
The odds ratio shows that the LGBTQ who took the placebo are 4.2 more times likely to get covid19 than those who took the vaccine.            
This means we reject the null hypothesis because the p-value of the fisher exact test is less than 0.05.          

### Conclusion
We conclude that the barchart shows the LGBTQ who took the vaccine were less likely to get infected with covid19 than the LGBTQ who took the placebo.            
The Numerical summary shows that that 97% of the LGBTQ who took the placebo got covid19 whereas 3% of the LGBTQ who took the vaccine got covid19.            
The inferential results show that the chi-squared test had a p-value of 0.000002266.       
This means we reject the null hypothesis that covid921 vaccine is not effective in the LGBTQ because the p-value is less than 0.05.     
The Fisher exact test had a p-value of 0.000000269.               
The odds ratio in the Fisher exact test shows that the LGBTQ who took the placebo are 1.8 more times likely to get covid19 than those who took the vaccine.          
This means we reject the null hypothesis that covid921 vaccine is not effective in the LGBTQ because the p-value is less than 0.05.          


## Drug Users
$H_0$:The covid921 vaccine is not effective in drug users 

$H_A$:The covid921 vaccine is effective in drug users  

### Methods
treatment is a categorical variable   
infected is a categorical variable   
So we use is cat~cat helper file   

### Descriptive Results 


#### Graphical Results
```{r}
DrugUser = filter(FinalData, DrugUser=="yes")
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

The barchart shows that Drug users who took the vaccine were more likely to get infected with covid19 than Drug users who took the placebo.    


#### Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical summary shows that 0.33% of Drug users who took the placebo got covid19 whereas 99.7% of Drug users who took the vaccine got covid19.      


#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test has a p-valye of 0.0000000001414.                    
This means we reject the null hypothesis because the p-value of the chi-squared is less than 0.05.          
The Fisher exact test had a p-value of 0.000000000003261.                
The odds ratio shows that Drug users who took the placebo are 0.078 times less likely to get covid19 than those who took the vaccine.                
This means we reject the null hypothesis because the p-value of the fisher exact test is less than 0.05.       

### Conclusion. 
We conclude that the barchart shows Drug users who took the vaccine were more likely to get infected with covid19 than Drug users who took the placebo.               
The Numerical summary shows that that 0.33% of Drug users who took the placebo got covid19 whereas 99.7% of Drug users who took the vaccine got covid19.                
The inferential results show that the chi-squared test had a p-value of 0.0000000001414.   
The Fisher exact test had a p-value of 0.000000000003261.                  
The odds ratio in the Fisher exact test shows that Drug users who took the placebo are 0.078 times less likely to get covid19 than those who took the vaccine.               
Although the graphical and numerical summaries show that Drug users who took the placebo were less likely to get covid19 than those who took the vaccine, we conclude that we reject the null hypotheis that covid921 vaccine is not effective in Drug users because the p-value of the chi-squared test and Fisher exact test were less than 0.05.   