---
title: "Covid19 vaccine study Final2024"
author: "Kaylie Fagoaga"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction
A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

##  Males
$H_0$:The covid921 vaccine is not effective in males 

$H_A$:The covid921 vaccine is effective in males 

###  Methods
treatment is a categorical variable   
infected is a categorical variable   
So we use is cat~cat helper file   
 
### Descriptive Results 


#### Graphical Results
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males, type="percent")

```

This barchart shows that males who took the vaccine were less likely to get infected with covid19 than the males who took the placebo.   


#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

The numerical summary shows that 74% of males who took the placebo got covid19 whereas 26% of males who took the vaccine got covid19.     

#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test has a p-valye of 0.00000000000000022.      
This means we reject the null hypothesis because the p-value of the chi-squared is less than 0.5.    
The Fisher exact test also had a p-value of 0.00000000000000022.      
The odds ratio shows that males who took the placebo are 2.8 more times likely to get covid19 than those who took the vaccine.      
This means we reject the null hypothesis because the p-value of the fisher exact test is less than 0.5.     

# Conclusion 
We conclude that the barchart shows males who took the vaccine were less likely to get infected with covid19 than the males who took the placebo.      
The Numerical summary shows that that 74% of males who took the placebo got covid19 whereas 26% of males who took the vaccine got covid19.   
The inferential results show that both the chi-squared test and Fisher exact test had a p-value of 0.00000000000000022.     
The odds ratio in the Fisher exact test shows that males who took the placebo are 2.8 more times likely to get covid19 than those who took the vaccine.  
This means we reject the null hypothesis that covid921 vaccine is not effective in males.

## Females
$H_0$:The covid921 vaccine is not effective in females 

$H_A$:The covid921 vaccine is effective in females

###  Methods
treatment is a categorical variable   
infected is a categorical variable   
So we use is cat~cat helper file  

### Descriptive Results 


#### Graphical Results
```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

The barchart shows that females who took the vaccine were less likely to get infected with covid19 than the females who took the placebo.   

#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected, data=females)
rowPerc(table1)
colPerc(table1)
```

The numerical summary shows that 65% of females who took the placebo got covid19 whereas 35% of females who took the vaccine got covid19.  


#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test has a p-valye of 0.00000001914.      
This means we reject the null hypothesis because the p-value of the chi-squared is less than 0.5.    
The Fisher exact test had a p-value of 0.00000001246.      
The odds ratio shows that females who took the placebo are 1.8 more times likely to get covid19 than those who took the vaccine.      
This means we reject the null hypothesis because the p-value of the fisher exact test is less than 0.5.    

# Conclusion 

## LGBTQ
$H_0$:The covid921 vaccine is not effective in the LGBTQ community 

$H_A$:The covid921 vaccine is effective in the LGBTQ community

###  Methods
treatment is a categorical variable   
infected is a categorical variable   
So we use is cat~cat helper file    
 
### Descriptive Results 


#### Graphical Results
```{r}
LGBTQ = filter(FinalData, LGBTQ=="gay")
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

#### Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
# Conclusion

## Drug Users
$H_0$:The covid921 vaccine is not effective in drug users 

$H_A$:The covid921 vaccine is effective in drug users  

### Methods
treatment is a categorical variable   
infected is a categorical variable   
So we use is cat~cat helper file   

### Descriptive Results 


#### Graphical Results
```{r}
DrugUser = filter(FinalData, DrugUser=="yes")
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

#### Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



# Conclusion. 